---
title: "Visulising edge-level brain data in R"
output: html_notebook
---

We will now move onto visualizing edge-level data. sometimes the result of our neuroimaging analyses result in egge-level values, in the form of a adjacency or connectivity matrix. The primary input is a n by n square-matrix, where n is the number of brain regions (aka nodes), and each element of a matrix related to a connectivity value (aka an edge). We will primarily be using the `brainconn` package which allows us to plot brain networks in 2d and interactive 3d. 

Lets first install `brainconn` and `pheatmap`

```{r, include=FALSE}
install.packages("remotes")
remotes::install_github("sidchop/brainconn")
install.packages("lattice")

```

`brainconn` comes with a few example connectivity matrices. We will first use the a unweighted (edges are either present or absent i.e. 1 or 0), and undirected (i.e. symmetric). Before we plot this network on a brain, lets just plot it at a square plot to get an idea of what kind of data we are working with. 
```{r}
library("brainconn")
x <- as.matrix(example_unweighted_undirected)
dim(x)
```

```{r}
library("lattice")
levelplot(x, main = "example_unweighted_undirected", ylab = "ROI", xlab = "ROI", scales=list(draw=FALSE))
```
Since the example matrix has 300 rois, which relate to the Schaefer 300 atlas, we will plot it using the same atlas. `brainconn` comes with many atlas, and it is simple to add your own [add link to atlas creation docs].

```{r}
x <- example_unweighted_undirected
brainconn(atlas ="schaefer300_n7", conmat=x, node.size = 3, view="ortho")
```

You can change the default view, node, edge and background properties as well as show all nodes of the atlas, rather than just the ones that have connected edges.  
```{r}
brainconn(atlas ="schaefer300_n7", 
          conmat=example_unweighted_undirected,
          view="left",
          node.size = 3, 
          node.color = "hotpink", 
          edge.width = 2, 
          edge.color="darkblue",
          edge.alpha = 0.8, 
          all.nodes = T, 
          show.legend = F)
```

If the connectivity matrix you input is not a binary matrix, i.e. the values are weighted, by default `braincon()` will modulate the `edge.width` by the weighted values. `scale.edge.width` can be used to scale the edge width.

```{r}
x <- example_weighted_undirected
brainconn(atlas ="schaefer300_n7", conmat=x, node.size = 5,view="top", scale.edge.width = c(1,3), background.alpha = 0.4, show.legend = F)
```
You can also color the edges by weight using the `edge.color.weighted` option and change the colour scale used using `scale_edge_colour_*` functions

```{r}
brainconn(atlas ="schaefer300_n7", 
          conmat=example_weighted_undirected, 
          node.size = 7,view="bottom", 
          edge.width = 2, 
          edge.color.weighted = T, 
          show.legend = T) +
  scale_edge_colour_gradient2(low='yellow', mid='orange', high='red')
```
If the input is a directed matrix (i.e. non-symmetrical), the edges will be displayed as directed arrows, with the `edge.width` and `edge.color` serving the same purpose:

```{r}
x <- example_unweighted_directed
brainconn(atlas ="schaefer300_n7", conmat=x, node.size = 4, view="right", edge.alpha = 0.6)
```

Weighted and directed matrix example:
```{r}
x <- example_weighted_directed
brainconn(atlas ="schaefer300_n7", conmat=x, view="front", edge.color.weighted=T) +
    scale_edge_colour_viridis() 
```


You can auto add `labels`, but this doesn't work to great if theirs a large number of edges. 
```{r}
brainconn(atlas ="schaefer300_n7", conmat=example_unweighted_undirected, labels = T, label.size = 2, node.size = 3)
```

If you want to weight the nodes by a feature such as degree, you can provide a vector the length of the number of ROIs in the parcellation to `node.size`: 

```{r}
x <- example_unweighted_undirected #convert connectivity matrix into an graph object.
d <- rowSums(x) #use igraph::degree to obtain a vector of nodal degree
d <- d[d != 0] #remove nodes with no edges 
brainconn(atlas ="schaefer300_n7", conmat=x, node.size = d)
```



## Basic use of `brainconn3D()`
Currently, `brainconn3D()` is only able to visualize binary undirected connectivity matrices. 

```{r,}
brainconn3D(atlas ="schaefer300_n7", conmat=example_unweighted_undirected, show.legend = F)

```

Modifiable features include `node.color`, `node.size`, `edge.size` & `edge.width`, brain surface `opacity` and `d.factor` (a multiplication factor to control the spread of nodes.)
```{r}
brainconn3D(atlas = "schaefer300_n7", conmat=example_unweighted_undirected, edge.width = 6, edge.color = "green", node.size = 8, node.color = "red", show.legend = F)

```


The `d.factor` can be adjusted for different purposes such as electrode height. The default is set to 1.05.
```{r}
brainconn3D(atlas ="schaefer300_n7", conmat=example_unweighted_undirected, edge.width = 3, edge.color = "brown", node.size = 3, d.factor = 1.3, all.nodes = T, opacity = 0.3, show.legend = F)

```