---
title: "Visulising ROI-level brain data in R"
output: html_notebook
---

We will now move onto visualising region level data. Often the result of our neuroimaging analyses are values assiged to spesifc brain regions. These values can be a test statistic (e.g. T-score), a effect size, a measure of cortical thickness or volume or anything at all. A great package for visualsing these values on brains is `ggseg` and `ggseg3D`. We will go though both below.

Lets first install the packages: 

#```{r, eval = FALSE}
#install.packages("ggseg")
#install.packages("ggseg3d")
#remotes::install_github("LCBC-UiO/ggsegSchaefer")
#```

```{r}
library("ggseg")
library("ggseg3d")
library("ggplot2")
library("ggsegSchaefer")
```

Calling the `ggseg` function will display the base graphic used by ggseg (left/right, medial/lateral images)

```{r}
ggseg()
```

There are many brain atleses avalibe in `ggseg`, but we will just use the Desikan-Killany cortical atlas for now, and colour seperatly each regions.
```{r}
ggseg(atlas = "dk", mapping=aes(fill=region))
```


Lets get rid of the legend, and stack the four images and change the thickness of the boundry around each region and add a title. Try fooling around with the colour and size parameters.

```{r}
ggseg(mapping=aes(fill=region), position="stacked",
        colour="black",size=.8,show.legend = F) +
  ggtitle("Desikan-Killany cortical atlas")
```


Now lets try colour the brain regious by some value which relates to our own input data. All atlases included with `ggseg` contain data.tables with all the regions included in the atlas. Lets first grab the all the regions in the atlas for both hemispheres. 

```{r}
base_atlas = as.data.frame(na.omit(cbind(dk$data$region, dk$data$hemi)))
colnames(base_atlas) <- c("region", "hemi")
head(base_atlas,5)
```

For now we are just going to genrate random numbers as input, but here you can use your own values.

```{r}
set.seed(3241)
brain_volume <- rnorm(dim(base_atlas)[1]) #here you would enter your own date that you wanted to plot
base_atlas <- cbind(base_atlas, brain_volume)
ggseg(base_atlas, mapping=aes(fill=brain_volume), position="stacked",
        colour="black",size=.8,show.legend = T) +
  ggtitle("Lets plot some values on the brain!")

```


Lets change the colour pallet used:

```{r}
ggseg(base_atlas, mapping=aes(fill=brain_volume), position="stacked",
        colour="darkblue",size=.8,show.legend = T) +
  ggtitle("Lets plot some values on the brain!") + scale_fill_gradientn(colours=c("dodgerblue4","light blue","firebrick","goldenrod"))
```

We can also select specific regions which are of interest to you. 

```{r}
regions <- data.frame(cbind(region=c("superior frontal","insula"),value=c(2.4,5.5)),hemi=c("right","right"),
                     stringsAsFactors=F)

ggseg(regions,mapping=aes(fill=as.numeric(value)), hemi= "right", colour="white",size=.3) 

```

`ggseg` also allows us to visualise subcortial structures.

```{r}
ggseg(atlas = "aseg")
```

Lets map some values onto subcortial structure. You can also remove regions we do not want by assigning them a `NA` value.

```{r}
regions = data.frame(cbind(region=c("hippocampus","pallidum",  "putamen"),
                           value=c(3,2,1)),
                     stringsAsFactors=F)

ggseg(atlas = "aseg",regions,mapping=aes(fill=as.numeric(value)), view = "axial") 

```


A sister package related to `ggseg`, called `ggseg3d` allows us to visualise our ROI level data in interactive 3D. We will briefly go though some of the basic funtionality. 

```{r}
ggseg3d(atlas = "schaefer7_3d")
```

Lets remove some of the axis labels and set the cammera to a angle we wont.

```{r}
plot <- ggseg3d(atlas = "schaefer7_3d")  
plot <- remove_axes(plot)
pan_camera(plot, "right medial") 
```


We can also view subcortical structures in 3D. 

```{r}
ggseg3d(atlas = "aseg_3d")  
```

Here is an example of function with advanced use of ggseg3d which selects 14 commonly used subcortial regions (7 each hemisphere), assigns a user-specifed value, and places the cammera at a nice angle. 

```{r}
make_ggseg3d <- function(attribute, 
                         colour.pal = c("light yellow",
                                        "orange",
                                        "red",
                                        "dark red"), 
                         hide.colourbar=FALSE, 
                         output.png=FALSE, 
                         file.name="ggseg_3d.png") {
# Inputs: 
#' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left-Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Right-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Right-Amygdala, Right-Accumbens-area

library(ggseg3d)
library(tidyr)
library(dplyr)
#remove(aseg_3d)
aseg_3d <- aseg_3d
aseg_3d <- tidyr::unnest(aseg_3d, cols = c(ggseg_3d))

attribute.ggseg3d <- c(rep(NA, 4), attribute[1:4], rep(NA, 3), attribute[5:7],
                       rep(NA, 5), attribute[8:14],  rep(NA, 6))

data <- dplyr::mutate(aseg_3d, attribute =  attribute.ggseg3d)
#remove NA regions
aseg_3d[which(is.na(data$attribute)),] <- NA
aseg_3d <- tidyr::drop_na(aseg_3d)

data[which(is.na(data$attribute)),] <- NA
data <- tidyr::drop_na(data)
data$attribute[data$attribute==0]<-NA #make 0 valus NA so they are set as grey in ggseg3d

scene=list(camera = list(eye = list(x = 0, y = 1, z = -2.25)),
           aspectratio = list(x=1.6,y=1.6,z=1.6))


plot <- ggseg3d::ggseg3d(.data = data, 
                 atlas = aseg_3d, 
                 colour = "attribute", 
                 text = "attribute",
                 palette = colour.pal)
plot <- remove_axes(plot)
plot <- plotly::layout(plot, 
                 scene = scene, 
                 plot_bgcolor  = "rgba(0, 0, 0, 0)",
                 paper_bgcolor = "rgba(0, 0, 0, 0)", 
                 width = 600, height = 600) 

if(hide.colourbar==TRUE){
  plot <- plotly::hide_colorbar(plot)
}

if( output.png==TRUE){
plotly::orca(plot, file =  file.name)
}

return(plot)

}
```

And here is an example use of this function:

```{r, eval = FALSE}
volume <- sample(1:20, 14) #randomly generated data for 14 structures
make_ggseg3d(volume, colour.pal = c("white", "firebrick"))
```

